Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-10-10 ms/review-responses.tex

\begin{point}{}
    I agree that most modern work describes structure as discrete populations connected by migration. However, some methods/studies have explicitly modeled spatial structure, e.g. especially in ecology or using methods like dadi (diffusion approximations). Highlighting some examples of previously identified structure not possible to infer without modeling geography would be helpful to contextualize this work.
\end{point}

\reply{
    \cjb{working on this}
}

\begin{point}{}
    There is some reference to spatial models using grids (e.g. Rousset 1997). Some additional discussion contextualizing more recent methods like EEMS that also construct demes and model migration through divergence between neighboring demes would be helpful and interesting.
\end{point}

\reply{
    Good point on EEMS. We have added the most recent EEMS paper to the citations on \llname{eemscitation}
}

@@

\begin{point}{}
    Demographic modeling. Both approaches tested, stairwayplot and SMC++, are most sensitive to older demographic events, and consequently are very noisy and underestimate effect population sizes, especially in smaller neighborhood sizes. Models that consider haplotype structure are much better suited to this time period. It would be helpful to either 1) discuss the varying time sensitivities of different classes of demographic inference methods and how spatial patterns of genetic variation would influence these inferences, or 2) apply a method of this class (many options, e.g. DoRIS, IBDNe, Tracts, Globetrotter, etc) and show how it performs.
\end{point}

\reply{
    We now discuss haplotype methods in the relevant discussion section \llname{ibdnesxn}. However though these methods should be more accurate for recent events it is not clear that this will improve performance per se. The dips in recent inferred Ne from stairwayplot are not just prediction noise, but actually reflect an underlying genealogy in which terminal branches are shorter than expected from a constant-size random-mating population (see e.g. figure \ref{fig:sumstats}A and \ref{fig:ibs}). The interpretation error is that these short branches are generated by spatial structure rather than changes in population size over time -- a point also made in the \citep{Mazet2015} paper we discuss in the introduction and discussion. 
}

@@

\begin{point}{}
    GWAS mixed models. To what extent can spatial signals (e.g. corner, patchy) be corrected with mixed models, e.g. with PCs and PC-adjusted GRM as in Conomos et al, 2016 using PC-AiR and PC-Relate)? Is patchiness related to dispersal? I'm curious how this relates to the predictive ability of GWAS phenotypes with some spatial association that may or may not be associated with environmental effects.
\end{point}

\reply{
\cjb{is this an ok response? two reviewers asked about mixed models but I feel like including them would require expanding the gwas section quite a bit and I'd rather now. But if either of you feel like it's critical I could probably get it done next week}
    Good question -- we are also interested to know how mixed models perform here, but think that to properly test that we would want to change our design to generate phenotypes from simulated genotypes. This would allow us to evaluate false-negatives in addition to false-positives. This is important because, if mixed models do provide stronger control for stratification they are also likely to remove true signals of causal SNPs if those SNPs covary with spatial structure. We now point to these methods explicitly in the discussion \llname{mixedmodelcallout}, but think that incorporating that study here would make this paper too long. We also think the PC results are still quite relevant as the method is still seen in many studies. 
}

@@

\begin{point}{}
    Code availability. This github link doesn't work, but is important to be able to evaluate for review: https://github.com/petrelharp/spaceness
\end{point}

\reply{
Apologies, it was accidentally set to private. The link should work now. 
}

@@

\begin{point}{}
    Definitions and interpretations. There are quite a large number of metrics discussed in Figure 3B, and it's a lot to take in. It might be helpful to have a table with a reminder of what the metric is, its interpretation, and how it is computed.
\end{point}

\reply{
    We have included a table describing the summary statistics in Figure \ref{table:sumstats}.
}
